Ricci-Ollivier curvature of two random walks on rooted phylogenetic subtree-prune-regraft graph

### 2015-02-21 curvature.tex
justification for lazy curvature definition

\begin{lemma}
	\label{lem:curvature_distance_bound}
	Let $T$ and $S$ be two trees. Then:
	$$ \frac{-2}{\dspr{T,S}} \le \curvature{T,S} \le \frac{2}{\dspr{T,S}} .$$
\end{lemma}
\begin{proof}
	Observe that the distance between neighbors of $T$ and $S$ is bounded between $\dspr{T,S} - 2$ and $\dspr{T,S} + 2$.
	For the curvature upper bound, we then have $\curvature{T,S} \le 1 - \frac{\dspr{T,S} + 2}{\dspr{T,S}} = \frac{2}{\dspr{T,S}}$.
	The lower bound follows similarly.
\end{proof}

@@

We next bound the effect of laziness on curvature.
The lazy random walk again selects a tree uniformly at random, as with the uniform random walk, but only accepts the move with probability $p$.
We denote the curvature of the lazy random walk between two trees $T$ and $S$ by $\curvature[\ell_p]{T,S}$.
For example, $\curvature[\ell_{\frac{1}{4}}]{T,S}$ describes the curvature of the lazy random walk that moves with probability $\frac{1}{4}$ and remains stationary with probability $\frac{3}{4}$.
For the lazy random walk, $m_T$ is now $T \cup N(T)$, with each neighbor assigned mass $\frac{p}{\degree{T}}$ and $T$ assigned the remaining $1 - p$ mass.
Note that $\curvature[\ell_p]{T, S} := \left(1 - \frac{W_1(m_T, m_S)}{d(x, y)}\right) / p$. \todo{EM: Do you have a good explanation for the lazy curvature definition?}

>>>>

\begin{lemma}
	\label{lem:curvature_distance_bound}
	Let $T$ and $S$ be two trees. Then:
	$$ \frac{-2}{\dspr{T,S}} \le \curvature{T,S} \le \frac{2}{\dspr{T,S}} .$$
\end{lemma}
\begin{proof}
	Observe that the distance between neighbors of $T$ and $S$ is bounded between $\dspr{T,S} - 2$ and $\dspr{T,S} + 2$.
	For the curvature upper bound, we then have $\curvature{T,S} \le 1 - \frac{\dspr{T,S} - 2}{\dspr{T,S}} = \frac{2}{\dspr{T,S}}$.
	The lower bound follows similarly.
\end{proof}

@@

We next bound the effect of laziness on curvature.
The lazy random walk again selects a tree uniformly at random, as with the uniform random walk, but only accepts the move with probability $p$.
We denote the curvature of the lazy random walk between two trees $T$ and $S$ by $\curvature[\ell_p]{T,S}$.
For example, $\curvature[\ell_{\frac{1}{4}}]{T,S}$ describes the curvature of the lazy random walk that moves with probability $\frac{1}{4}$ and remains stationary with probability $\frac{3}{4}$.
For the lazy random walk, $m_T$ is now $T \cup N(T)$, with each neighbor assigned mass $\frac{p}{\degree{T}}$ and $T$ assigned the remaining $1 - p$ mass.
Note that Ollivier defined $\curvature[\ell_p]{T, S} := \left(1 - \frac{W_1(m_T, m_S)}{d(x, y)}\right) / p$~\cite{Ollivier2009-bw}.
This definition of curvature is invariant of $p$ for small enough $p$ and was used by Ollivier to avoid parity problems on graphs where the uniform random walk is periodic.
As we now prove, the curvature of the lazy random walk is closely related to that of the uniform random walk.
